---
title: "Assigment"
author: "Kyaw Min Khaing"
date: "2025-02-11"
output: html_document
---

```{r setup, include=FALSE}
library(pacman)
p_load(tidyverse, rstatix, ggplot2, knitr, kableExtra)
```


```{r}
knitr::opts_chunk$set(
  echo = TRUE,    
  warning = FALSE, 
  message = FALSE, 
  fig.width = 6,   
  fig.height = 4   
)
```

## Hypothesis Statements

- **Null Hypothesis (H₀)**: The median ranks of math ability are the same across the three instructional methods.
- **Alternative Hypothesis (H₁)**: At least one instructional method leads to a different median rank in math ability.

```{r}
data <- tibble(
  Student = 1:15,
  Class = rep(c("Class 1", "Class 2", "Class 3"), each = 5),
  Score = c(85, 90, 88, 92, 89, 78, 83, 75, 80, 76, 72, 79, 77, 74, 78)
)
kable(data, caption = "student scores")
```
student score
```{r}
data_long <- data %>% pivot_longer(cols = Score, names_to = "Variable", values_to = "Value")

ranked_data <- data_long |> 
  mutate(Rank = rank(Value), 
         Tie_Case = duplicated(Value) | duplicated(Value, fromLast = TRUE)) |> 
  arrange(Rank)  # Rank ạảặổẹ ấẼộ ặếặỗ

kable(ranked_data, caption = "Sorting, Ranking and Tie")
```
total rank
```{r}
rank_sums <- ranked_data %>% 
  group_by(Class) %>% 
  summarise(Rank_Sum = sum(Rank), n = n())
kable(rank_sums, caption = "Total Ranks per Class")
```


```{r}
N <- nrow(ranked_data)

cat("Total number of observation (N):", N, "\n")
```

```{r}
k <- n_distinct(ranked_data$Class)
cat(" Number of category (N):", k, "\n")
```

## Compute the Kruskal-Wallis H-test statistic using the formula:

$$
H = \frac{12}{N(N+1)} \sum \frac{R_i^2}{n_i} - 3(N+1)$$


Where: - $N$ is the total number of observations. - $R_i$ is the rank sum for each group. - $n_i$ is the number of observations in each group.
```{r}
df <- k - 1

H <- 12 / (N * (N + 1)) * sum((rank_sums$Rank_Sum^2) / rank_sums$n) - 3 * (N + 1)
p_value <- pchisq(H, df, lower.tail = FALSE)

cat("Kruskal-Wallis H statistic:", round(H, 4), "\n")
cat("Degrees of Freedom:", df, "\n")
cat("p-value:", round(p_value, 6), "\n")
if (p_value < 0.05) {
  cat("The result is significant (p < 0.05). At least one group differs.\n")
} else {
  cat("The result is not significant (p >= 0.05). Groups are similar.\n")
}
```

```{r}
ties_table <- ranked_data |> 
  filter(Tie_Case) |> 
  count(Rank)

# Print the updated ties table
kable(ties_table, caption = "Number of Tie")
```
# Compute C_H correction factor

## To Compute the Test Statistic doing Ties Correction (C_H)

The ties correction is applied to adjust for tied ranks. The formula for the ties correction $C_H$ is:

$$
C_H = 1 - \frac{\sum{(T^3 - Ti)}}{N^3 - N}
$$

Where: - $T$ is the number of tied ranks in each group of ties. - $N$ is the total number of observations.

```{r}
C_H <- 1 - sum((ties_table$n^3 - ties_table$n) / ((nrow(ranked_data)^3) - nrow(ranked_data)))

# Print C_H
cat("Ties Correction Factor (C_H):", round(C_H, 4), "\n")
```


```{r}
H_corrected <- H / C_H 
cat("Corrected H Statistic:", round(H_corrected, 4), "\n")
```

```{r}
pairwise_results <- ranked_data %>% pairwise_wilcox_test(Rank ~ Class, p.adjust.method = "bonferroni")
kable(pairwise_results, caption = "Pairwise Wilcoxon Rank Sum Test Results")
```


```{r}
ggplot(ranked_data, aes(x = Class, y = Rank, fill = Class)) +
  geom_boxplot(outlier.color = "red", alpha = 0.7) +
  geom_jitter(aes(color = Class), width = 0.2, size = 2, alpha = 0.7) +
  labs(title = "Math Ability Rank Distribution by Class",
       x = "Instructional Method",
       y = "Rank (Lower is Better)") +
  theme_minimal() +
  scale_fill_manual(values = c("Class 1" = "blue", "Class 2" = "green", "Class 3" = "purple")) +
  scale_color_manual(values = c("Class 1" = "blue", "Class 2" = "green", "Class 3" = "purple")) +
  theme(legend.position = "none")
```
## Conclusion
```{r}
if (p_value < 0.05) {
  cat("The result is significant (p < 0.05). At least one group differs in math ability rankings.\n")
} else {
  cat("The result is not significant (p >= 0.05). No significant difference in math ability rankings across classes.\n")
}
```
