---
title: "ForexamSample Spearman Rank-Order Correlation (Small_Data_Samples _with _Ties)"
author: "Kyaw Min Khaing"
date: "2025-01-29"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Introduction

This analysis applies the **Spearman Rank-Order Correlation** to examine the relationship between **Number of Visits** and **Mean Heart Rate** for a small dataset with ties.

## Step 1: State the Hypotheses

-   **Null Hypothesis (H₀):** ρs = 0 There is no significant correlation between the number of visits and the mean heart rate.
-   **Alternative Hypothesis (H₁):**ρs ≠ 0 There is a significant correlation between the number of visits and the mean heart rate.

## Step 2: Set the Level of Significance

We set **α = 0.05,**there is a 95% chance ,that any observed statistical difference will be real and not due to chance.

```{r cars}
data <- tibble(
  Participant = 1:13,
  Visits = c(5, 12, 7, 14, 3, 8, 15, 12, 2, 16, 12, 7, 17),
  HeartRate = c(96, 63, 78, 66, 79, 95, 67, 64, 99, 62, 65, 76, 61)
)
print(data)
```

## Rank the data with ties correction

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ranked_data <- data %>%
  mutate(
    Rank_Visits = rank(Visits, ties.method = "average"),
    Rank_HeartRate = rank(HeartRate, ties.method = "average"),
    D = Rank_Visits - Rank_HeartRate,
    D_squared = D^2,
    Ties_Visits = ifelse(duplicated(Rank_Visits) | duplicated(Rank_Visits, fromLast = TRUE), TRUE, FALSE),
    Ties_HeartRate = ifelse(duplicated(Rank_HeartRate) | duplicated(Rank_HeartRate, fromLast = TRUE), TRUE, FALSE)
  )

# Count tie groups for correction
T_visits <- sum(table(ranked_data$Rank_Visits)^3 - table(ranked_data$Rank_Visits)) / 12
T_heart_rate <- sum(table(ranked_data$Rank_HeartRate)^3 - table(ranked_data$Rank_HeartRate)) / 12
ranked_data <- ranked_data %>%
  arrange(Rank_Visits, Rank_HeartRate)
print(ranked_data)

```

## Step 4: Compute Spearman's Rank Correlation Coefficient with Ties Correction

The formula for Spearman's Rank Correlation Coefficient with ties correction is:

$$
 r_s = \frac{(n^3 - n) - 6 \sum D^2 - (T_x + T_y)/2}{\sqrt{(n^3 - n)^2 - (T_x + T_y)(n^3 - n) + T_x T_y}}
$$ where: - $D$ is the difference between ranks, - $n$ is the number of observations, - $T_x$ and $T_y$ are tie correction factors for Visits and Heart Rate.

```{r}
n <- nrow(ranked_data)
D_squared_sum <- sum(ranked_data$D_squared)

numerator <- (n^3 - n) - 6 * D_squared_sum - (T_visits + T_heart_rate) / 2
denominator <- sqrt((n^3 - n)^2 - (T_visits + T_heart_rate) * (n^3 - n) + T_visits * T_heart_rate)

r_s <- numerator / denominator

cat("Spearman Rank Correlation Coefficient (ρ):", round(r_s, 4), "\n")
```

## Step 5: Determine the Critical Value

For **n = 13** and **α = 0.05 (two-tailed test)**, the critical value from the **Spearman correlation table** is **±0.552**.

## Step 6: Compare the Obtained Value with the Critical Value

```{r}
if (abs(r_s) >= 0.552) {
  cat("The correlation is statistically significant. We reject H₀ and conclude there is a significant correlation.\n")
} else {
  cat("The correlation is not statistically significant. We fail to reject H₀.\n")
}
```

## Step 7: Interpret the Results

If **\|ρ\| ≥ 0.552**, the result is significant.

## Step 8: Reporting the Results

> *A Spearman rank-order correlation was conducted to determine the relationship between the number of visits and mean heart rate. The analysis yielded ρ = `r round(r_s, 4)`. Since the obtained ρ is `r ifelse(abs(r_s) >= 0.552, "greater", "less")` than the critical value of 0.552, we `r ifelse(abs(r_s) >= 0.552, "reject", "fail to reject")` the null hypothesis. This suggests that there `r ifelse(abs(r_s) >= 0.552, "is", "is no")` significant correlation between the two variables.*

## Step 9: Data Visualization

To better understand the ranked data distribution, we use `pivot_longer` to reshape the data and visualize it using `ggplot2`.

```{r}
# Convert data to long format for visualization
data_long <- ranked_data %>%
  pivot_longer(cols = c(Rank_Visits, Rank_HeartRate), names_to = "Variable", values_to = "Rank")

# Create visualization
rank_plot <- ggplot(data_long, aes(x = Variable, y = Rank, fill = Variable)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +
  labs(
    title = "Rank Distribution of Visits and Heart Rate",
    x = "Variable",
    y = "Rank"
  ) +
  theme_minimal()

# Print the plot
print(rank_plot)
```
