---
title: "ForexamSample Spearman Rank-Order Correlation (Small_Data_Samples _without
  _Ties)"
author: "Kyaw Min Khaing"
date: "2025-01-29"
output:
  word_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse,knitr)
```

## Introduction

This analysis applies the **Spearman Rank-Order Correlation** to examine
the relationship between **Number of Visits** and **Mean Heart Rate**
for a small dataset without ties.

## Step 1: State the Hypotheses

-   **Null Hypothesis (H₀):** There is no significant correlation
    between the number of visits and the mean heart rate.
-   **Alternative Hypothesis (H₁):** There is a significant correlation
    between the number of visits and the mean heart rate. \## Step 2:
    Set the Level of Significance We set **α = 0.05** as the
    significance level. \## Step 3: Load and Prepare Data

```{r cars}
data <- tibble(
  Participant = 1:8,
  Visits = c(5, 12, 7, 14, 2, 8, 15, 17),
  HeartRate = c(100, 89, 78, 66, 77, 103, 67, 63)
)
kable(data, caption = "Heart Rate")
```

## Rank the data

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ranked_data <- data %>%
  mutate(
    Rank_Visits = rank(Visits, ties.method = "average"),
    Rank_HeartRate = rank(HeartRate, ties.method = "average"),
    D = Rank_HeartRate - Rank_Visits   ,
    D_squared = D^2,
    Ties_Visits = ifelse(duplicated(Rank_Visits) | duplicated(Rank_Visits, fromLast = TRUE), TRUE, FALSE),
    Ties_HeartRate = ifelse(duplicated(Rank_HeartRate) | duplicated(Rank_HeartRate, fromLast = TRUE), TRUE, FALSE)
  )
kable(ranked_data, caption = "Heart Rate D square and Tie")
```

## Step 4: Compute Spearman's Rank Correlation Coefficient

The formula for Spearman's Rank Correlation Coefficient is:

$$
  r_s = 1 - \frac{6 \sum D^2}{n(n^2 - 1)}
$$
where: - $D$ is the difference between ranks, - $n$ is the number of
observations.

```{r}
n <- nrow(ranked_data)
D_squared_sum <- sum(ranked_data$D_squared)

r_s <- 1 - (6 * D_squared_sum) / (n * (n^2 - 1))

cat("Spearman Rank Correlation Coefficient (ρ):", round(r_s, 4), "\n")
```
```{r}
cor_test_result <- data %>%
  cor_test(Visits, HeartRate, method = "spearman")
kable(cor_test_result, caption = "cor")
```


## Step 5: Determine the Critical Value

For **n = 8** and **α = 0.05 (two-tailed test)**, the critical value
from the **Spearman correlation table** is **±0.738**.

## Step 6: Compare the Obtained Value with the Critical Value

```{r}
if (abs(r_s) >= 0.738) {
  cat("The correlation is statistically significant. We reject H₀ and conclude there is a significant correlation.\n")
} else {
  cat("The correlation is not statistically significant. We fail to reject H₀.\n")
}
```

## Step 7: Interpret the Results

If **\|ρ\| ≥ 0.738**, the result is significant.

## Step 8: Reporting the Results

> *A Spearman rank-order correlation was conducted to determine the
> relationship between the number of visits and mean heart rate. The
> analysis yielded ρ = `r round(r_s, 4)`. Since the obtained ρ is
> `r ifelse(abs(r_s) >= 0.738, "greater", "less")` than the critical
> value of 0.738, we
> `r ifelse(abs(r_s) >= 0.738, "reject", "fail to reject")` the null
> hypothesis. This suggests that there
> `r ifelse(abs(r_s) >= 0.738, "is", "is no")` significant correlation
> between the two variables.*

```{r}
# Convert data to long format for visualization
ggplot(data, aes(x = Visits, y = HeartRate)) +
  geom_point(size = 4, color = "blue") +  # Data points များကို ပြသပါ
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Trend line ထည့်ပါ
  labs(
    title = "Scatterplot of Visits vs HeartRate",
    x = "Number of Visits",
    y = "Heart Rate"
  ) +
  theme_minimal() 
```
