---
title: "Forexam122"
author: "Kyaw Min Khaing"
date: "2025-01-28"
output:
  word_document: default
  html_document: default
---
## State the Null and Research Hypotheses

The null hypothesis is
HO: θL = θM = θH
The research hypothesis is
HA: There is a tendency for self-conidence to rank systematically higher or 
lower for at least one level of social interaction when compared with the 
other levels.
## Set the Level of Risk (or the Level of Signiicance) Associated with the Null Hypothesis
α = 0.05
95% chance that any observed statistical difference will be real and not due to chance
## need library this package
```{r setup, include=FALSE}
library(pacman)

p_load(tidyverse,rstatix,ggsignif)
```
## Data Preparation
The table below shows the sample data for three groups (High, Medium, Low)

```{r data-preparation}
data <- data.frame(
  High = c(21, 23, 18, 12, 19, 20),
  Medium = c(19, 5, 10, 11, 9, NA),
  Low = c(7, 8, 15, 3, 6, 4)
)
print(data)
```
# Convert to long format for analysis
```{r}
data_long <- data %>%
  pivot_longer(cols = everything(), names_to = "Group", values_to = "Value") %>%
  drop_na()
print(data_long)
# Print the data
```
## Ranking Data and Handling Ties
The Kruskal-Wallis test uses ranks rather than raw data values. We rank all values, including ties.
```{r}
ranked_data <- data_long %>%
  mutate(Rank = rank(Value))

# Display ranked data
print(ranked_data)
```
##  To Compute the Test Statistic doing Group Ranks and Sums
We calculate the rank sum for each group.
```{r}
# Calculate rank sums
rank_sums <- ranked_data %>%
  group_by(Group) %>%
  summarise(
    n = n(),
    Rank_Sum = sum(Rank)
  )

# Print rank sums
print(rank_sums)
```
## To Compute the Test Statistic doing Ties Correction (C_H)
The ties correction is applied to adjust for tied ranks.
The formula for the ties correction \( C_H \) is:

\[
C_H = 1 - \frac{\sum{(T^3 - Ti)}}{N^3 - N}
\]

Where:
- \( T \) is the number of tied ranks in each group of ties.
- \( N \) is the total number of observations.

```{r}
ties_table <- ranked_data %>%
  count(Rank)

C_H <- 1 - sum((ties_table$n^3 - ties_table$n) / ((nrow(ranked_data)^3) - nrow(ranked_data)))

# Print ties correction factor
cat("Ties Correction Factor (C_H):", C_H, "\n")
```
## Compute the H-Test Statistic
We calculate the Kruskal-Wallis H-test statistic using the formula:
\[
H = \frac{12}{N(N+1)} \sum \frac{R_i^2}{n_i} - 3(N+1)\]

Where:
- \(N\) is the total number of observations.
- \(R_i\) is the rank sum for each group.
- \(n_i\) is the number of observations in each group.
```{r}
# Total number of observations
N <- nrow(ranked_data)
# Calculate H statistic
H <- 12 / (N * (N + 1)) * sum((rank_sums$Rank_Sum^2) / rank_sums$n) - 3 * (N + 1)
H_corrected <- H / C_H  # Apply ties correction
# Print H statistic
cat("H-Test Statistic (Corrected):", round(H_corrected, 4), "\n")
```
## Significance Test
Compare the H statistic to the Chi-Square distribution with \(k-1\) degrees of freedom (where \(k\) is the number of groups) to determine significance.

```{r}
# Degrees of freedom
k <- ranked_data %>%
  summarise(num_groups = n_distinct(Group)) %>%
  pull(num_groups)
df <- k - 1

# Critical value and p-value
alpha <- 0.05
p_value <- pchisq(H_corrected, df, lower.tail = FALSE)

# Report significance
cat("Degrees of Freedom (df):", df, "\n")
cat("p-value for Kruskal-Wallis H-Test:", round(p_value, 4), "\n")
if (p_value < alpha) {
  cat("The result is significant at α =", alpha, "\n")
} else {
  cat("The result is not significant at α =", alpha, "\n")
}
```

```{r}
# Pairwise Mann-Whitney U tests
pairwise_results <- ranked_data %>%
  pairwise_wilcox_test(
    Value ~ Group,
    p.adjust.method = "bonferroni"
  )

# Display results
print(pairwise_results)
```
 Final Report
```{r}
cat("### Final Report\n")
cat("Three groups were analyzed: High (n = 6), Medium (n = 5), and Low (n = 6).\n")
cat("The corrected Kruskal-Wallis H statistic was", round(H_corrected, 4), "with", df, "degrees of freedom.\n")
cat("The p-value was", round(p_value, 4), "at α =", alpha, ".\n")
if (p_value < alpha) {
  cat("We reject the null hypothesis and conclude that at least one group differs significantly.\n")
} else {
  cat("We fail to reject the null hypothesis, suggesting no significant differences between the groups.\n")
}

# Pairwise significance summary
cat("\n### Pairwise Comparisons (Mann-Whitney U Test):\n")
cat("Results adjusted using Bonferroni correction:\n")
```
## Visualization
Visualize the ranked data for each group using a boxplot.
```{r}
pairwise_results <- data_long %>%
  pairwise_wilcox_test(Value ~ Group, p.adjust.method = "bonferroni")

# Create the rank plot with significance brackets
rank_plot <- ggplot(ranked_data, aes(x = Group, y = Rank, fill = Group)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  labs(
    title = "Rank Distribution by Group",
    x = "Group",
    y = "Rank"
  ) +
  theme_minimal() +
  geom_signif(
    comparisons = list(
      c("High", "Medium"),
      c("High", "Low"),
      c("Medium", "Low")
    ),
    map_signif_level = TRUE,
    test = "wilcox.test",  # Perform Mann-Whitney U test directly
    step_increase = 0.1   # Adjust spacing to avoid overlapping
  )

# Print the plot
print(rank_plot)
```


