---
title: "Mortality Rate Comparison of Country A and B"
author: "Kyaw Min Khaing"
date: "2025-02-04"
output: html_document
---

## Mortality Rate Comparison
The mortality conditions of **Country A** and **Country B** are calculated using:

### **Crude Death Rate (CDR) for Country A**
$$
CDR(A) = \frac{\sum M_x(A) P_x(A)}{\sum P_x(A)} \times 1000
$$

### **Crude Death Rate (CDR) for Country B**
$$
CDR(B) = \frac{\sum M_x(B) P_x(B)}{\sum P_x(B)} \times 1000
$$

### **Standardized Death Rate (SDR) for Country A (using Country B's age structure)**
$$
SDR(A/B) = \frac{\sum M_x(A) P_x(B)}{\sum P_x(B)} \times 1000
$$

---

```{r}
library(tidyverse)
```

## **Data and Calculation in R**
```{r}
# Given Data
mortality_data <- tibble(
  Age_Group = c("0-4", "5-14", "15-24", "25-64", "65+"),
  Px_A = c(660, 1200, 1000, 1975, 165),
  Mx_A = c(0.025, 0.004, 0.005, 0.006, 0.050),
  Px_B = c(595, 1275, 1340, 5350, 1440),
  Mx_B = c(0.020, 0.004, 0.006, 0.006, 0.030),
)
print(mortality_data)
```
# Compute Required Values
```{r}
mortality_data <- mortality_data %>%
  mutate(
    MxA_PxA = Mx_A * Px_A,
    MxB_PxB = Mx_B * Px_B,
    MxA_PxB = Mx_A * Px_B,
    MxB_PxA = Mx_B * Px_A
  )
print(mortality_data)
```
# Summarizing Values
```{r}
total_PxA <- sum(mortality_data$Px_A)
total_PxB <- sum(mortality_data$Px_B)
total_MxA_PxA <- sum(mortality_data$MxA_PxA)
total_MxB_PxB <- sum(mortality_data$MxB_PxB)
total_MxA_PxB <- sum(mortality_data$MxA_PxB)
total_MxB_PxA <- sum(mortality_data$MxB_PxA)
```
# CDR & SDR Calculations
```{r}
CDR_A <- (total_MxA_PxA / total_PxA) * 1000
CDR_B <- (total_MxB_PxB / total_PxB) * 1000
SDR_A_B <- (total_MxA_PxB / total_PxB) * 1000
SDR_B_A <- (total_MxB_PxA / total_PxA) * 1000
```
# Display Results in a tibble
```{r}
results <- tibble(
  Metric = c("CDR(A)", "CDR(B)", "SDR(A/B)", "SDR(B/A)"),
  Value = c(round(CDR_A, 2), round(CDR_B, 2), round(SDR_A_B, 2), round(SDR_B_A, 2))
)

mortality_data
results
```
Visualization of Mortality Rates using ggplot
```{r}
# Plot Mortality Rates
ggplot(results, aes(x = Metric, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(
    title = "Comparison of CDR(A), CDR(B), SDR(A/B), and SDR(B/A)",
    x = "Metric",
    y = "Value (per 1000 population)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(Value, 2)), vjust = -0.5)
```


