---
title: "Demographic Analysis of Mauritius (1989)"
author: "Kyaw Min Khaing"
date: "2025-02-05"
output: word_document
---
## **Introduction**
This analysis examines the population dynamics of Mauritius in **1989**, focusing on:  
1. **Natural Increase (NI)**  
2. **Crude Rate of Natural Increase (CRNI)**  
3. **Vital Index (VI)**  
---
## **Formulae for Calculations**
1. **Natural Increase (NI):**
   $$
   NI = B - D
   $$

2. **Crude Rate of Natural Increase (CRNI):**
   $$
   CRNI = \frac{(B - D)}{P} \times 1000
   $$

3. **Vital Index (VI):**
   $$
   VI = \frac{B}{D} \times 100
   $$

where:  
- \( B = 20,875 \) (Total Births)  
- \( D = 6,946 \) (Total Deaths)  
- \( P = 1,064,000 \) (Mid-Year Population)  

---

## **Data Representation**

```{r setup, include=FALSE}
library(tidyverse)
```

# Given Data

```{r}
mauritius_data <- tibble(
  Year = 1989,
  Births = 20875,
  Deaths = 6946,
  Population = 1064000,
  Natural_Increase = 20875 - 6946, # NI Calculation
  CRNI = ((20875 - 6946) / 1064000) * 1000, # CRNI Calculation
  Vital_Index = (20875 / 6946) * 100 # VI Calculation
)

# Display the tibble table
mauritius_data

```

Visualization: Births, Deaths, and Natural Increase

```{r pressure, echo=FALSE}
ggplot(mauritius_data, aes(x = Year)) +
  geom_col(aes(y = Births), fill = "blue", alpha = 0.6) +
  geom_col(aes(y = Deaths), fill = "red", alpha = 0.6) +
  geom_text(aes(y = Births, label = Births), vjust = -0.5, size = 5) +
  geom_text(aes(y = Deaths, label = Deaths), vjust = 1.5, size = 5, color = "white") +
  labs(title = "Births vs. Deaths in Mauritius (1989)",
       x = "Year",
       y = "Number of People",
       fill = "Legend") +
  theme_minimal()
```
Visualization: Crude Rate of Natural Increase
```{r}
ggplot(mauritius_data, aes(x = "", y = CRNI, fill = "CRNI per 1000")) +
  geom_bar(stat = "identity", width = 0.5) +
  coord_polar("y", start = 0) +
  labs(title = "Crude Rate of Natural Increase (1989)",
       subtitle = "Mauritius",
       fill = "Rate per 1000") +
  theme_minimal()
```
Visualization: Vital Index
```{r}
ggplot(mauritius_data, aes(x = "", y = Vital_Index, fill = "Vital Index")) +
  geom_bar(stat = "identity", width = 0.5) +
  coord_polar("y", start = 0) +
  labs(title = "Vital Index of Mauritius (1989)",
       subtitle = "Births per 100 Deaths",
       fill = "Index") +
  theme_minimal()
```

