---
title: "Maternal Mortality Ratio (MMR) Calculation & Visualization"
author: "Kyaw Min Khaing"
date: "2025-02-04"
output: word_document
---

```{r setup, include=FALSE}
library(pacman)
p_load(tidyverse,ggplot2,kableExtra,latex2exp)
```

Given Data

```{r}
# Define given values
P <- 350000  # Total Population
B <- 75000   # Total Live Births
D_p <- 245   # Pregnancy-Related Deaths
```

# Compute MMRo (Maternal Mortality Ratio)
## Maternal Mortality Ratio (MMRo)
The formula for the Maternal Mortality Ratio (MMRo) is:

$$
MMR_o = \left( \frac{D_p}{B} \right) \times 1000
$$

Where:  
- \( D_p \) = Number of pregnancy-related deaths  
- \( B \) = Total live births 
You can also embed plots, for example:

```{r}
MMRo <- (D_p / B) * 1000
print(MMRo)
```

# Compute Maternal Mortality (MM)
## Maternal Mortality (MM)
The formula for Maternal Mortality is:

$$
Maternal\ Mortality = \frac{MMR_o \times P}{1000}
$$

Where:  
- \( P \) = Total population  
- \( MMR_o \) = Maternal Mortality Ratio  
```{r}
Maternal_Mortality <- (MMRo * P) / 1000
print(Maternal_Mortality)
```

# Create a tibble table

```{r}
mmr_data <- tibble(
  Metric = c("Total Population", "Total Live Births", "Pregnancy-Related Deaths", "Maternal Mortality Ratio (MMRo)", "Maternal Mortality"),
  Value = c(P, B, D_p, round(MMRo, 2), round(Maternal_Mortality))
)
print(mmr_data)

```

# ✅ Ensure correct usage of dplyr filter()

```{r}
mmr_data_filtered <- mmr_data %>% 
  filter(Metric == "Maternal Mortality Ratio (MMRo)")
print(mmr_data_filtered)
```


Maternal Mortality Bar Chart
```{r}
ggplot(mmr_data_filtered , aes(x = Metric, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", width = 0.4) +
  labs(
    title = "Maternal Mortality Ratio (MMRo)",
    x = "Metric",
    y = "Value (per 1000 Live Births)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(Value, 2)), vjust = -0.5)
```



