---
title: "MAS122Cause-Specific Death Ratio & Rate in Egypt (1987)"
author: "Kyaw Min Khaing"
date: "2025-02-05"
output:
  word_document: default
  html_document: default
---

## **Introduction**

This analysis calculates the **Cause-Specific Death Ratio (CSDR)** and **Cause-Specific Death Rate (CSDRate)** for **hyperplasia of the prostate** in Egypt in 1987.

---
## **Formulae for Calculation**
$$
CSDR = \left( \frac{D_c}{D} \right) \times 100
$$

$$
CSDRate = \left( \frac{D_c}{P} \right) \times 100,000
$$

where:  
- \( D_c = 3,319 \) (Deaths due to hyperplasia of the prostate)  
- \( D = 466,161 \) (Total deaths)  
- \( P = 49,089,000 \) (Mid-year population)  
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Given Data

```{r cars}
death_data <- tibble(
  Year = 1987,
  Total_Deaths = 466161,
  Deaths_Hyperplasia = 3319,
  Population = 49089000,
  CSDR = (3319 / 466161) * 100,  # Cause-Specific Death Ratio
  CSDRate = (3319 / 49089000) * 100000  # Cause-Specific Death Rate per 100,000
)

# Display the tibble table
death_data
```

## Visualization: Cause-Specific Death Ratio

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ggplot(death_data, aes(x = "", y = CSDR, fill = "Hyperplasia of Prostate")) +
  geom_bar(stat = "identity", width = 0.5) +
  coord_polar("y", start = 0) +
  labs(title = "Cause-Specific Death Ratio (Hyperplasia of Prostate)",
       subtitle = "Egypt, 1987",
       fill = "Cause") +
  theme_minimal()
```

Visualization: Cause-Specific Death Rate

```{r}
ggplot(death_data, aes(x = Year, y = CSDRate)) +
  geom_col(fill = "darkred") +
  geom_text(aes(label = round(CSDRate, 2)), vjust = -0.5, size = 5) +
  labs(title = "Cause-Specific Death Rate per 100,000 Population",
       x = "Year",
       y = "Death Rate") +
  theme_minimal()
```
