---
title: "Fertility Measures Calculation"
author: "Kyaw Min Khaing"
date: "`r Sys.Date()`"
output: word_document
---

## Total Fertility Rate (TFR)

$$
TFR = 5 \sum f_x
$$

-   The TFR measures the total number of children a woman would have if age-specific fertility rates (ASFR) remain constant.
-   **Result:** $TFR = 3$ → Each woman would have about 3 births on average.

## General Fertility Rate (GFR)

$$
GFR = \frac{B}{F_{15-49}} \times 1000
$$

where:\
- $B$ = Total number of live births\
- $F_{15-49}$ = Number of females aged 15-49

-   The GFR measures the number of live births per 1000 women in the reproductive age group (15-49 years).
-   **Result:** $GFR = 102$ → There were 102 live births per 1000 women aged 15-49.

## Gross Reproduction Rate (GRR)

$$
GRR = 5 \sum f_x^*
$$

where:\
- $f_x^*$ = Age-specific fertility rate (ASFR) for female births only

-   The GRR measures the number of daughters a woman would produce if age-specific fertility rates remain constant.
-   **Result:** $GRR = 2$ → Each woman would produce 2 daughters on average.

```{r setup, include=FALSE}
library(tidyverse)
```

## Fertility Data Table

```{r}
# Data entry using tibble
fertility_data <- tibble(
  Age_Group = c("15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "Total"),
  Females_000 = c(465, 432, 360, 321, 289, 227, 170, 2264),
  Live_Births_000 = c(26, 65, 61, 42, 21, 10, 7, 232),
  Female_Live_Births_000 = c(13, 42, 35, 28, 12, 6, 3, 139),
  ASFR = c(0.0559, 0.1505, 0.1694, 0.1308, 0.0727, 0.0441, 0.0412, 0.6646),
  ASFR_Female = c(0.0279, 0.0972, 0.0972, 0.0872, 0.0415, 0.0264, 0.0176, 0.3950)
)

# Display tibble table
fertility_data
```

## Including Plots

You can also embed plots, for example: Total Fertility Rate (TFR)

```{r pressure, echo=FALSE}
# Compute TFR
TFR <- 5 * sum(fertility_data$ASFR[1:7])
TFR
```

General Fertility Rate (GFR) The General Fertility Rate (GFR) is calculated as:

```{r}
# Compute GFR
GFR <- (fertility_data$Live_Births_000[8] / fertility_data$Females_000[8]) * 1000
GFR
```

Gross Reproduction Rate (GRR)

```{r}
# Compute GRR
GRR <- 5 * sum(fertility_data$ASFR_Female[1:7])
GRR
```

Visualization of ASFR

```{r}
ggplot(fertility_data, aes(x = Age_Group, y = ASFR)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Age-Specific Fertility Rate (ASFR)",
       x = "Age Group",
       y = "ASFR")
```

Comparison of TFR, GFR, and GRR

```{r}
fertility_measures <- tibble(
  Measure = c("TFR", "GFR", "GRR"),
  Value = c(TFR, GFR, GRR)
)

ggplot(fertility_measures, aes(x = Measure, y = Value, fill = Measure)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Comparison of Fertility Measures",
       x = "Measure",
       y = "Value")
```
