---
title: "Two-Sample T-Test for Engine Mounting Time"
author: "Kyaw Min Khaing"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Define Hypotheses and Identify the Claim
We want to determine if there is a significant difference in engine mounting times using the Welles method (W) and the Atkins method (A).

Null Hypothesis (H₀):  (The mean mounting times are equal)
Alternative Hypothesis (H₁):(The mean mounting times are different)
This is a two-tailed test at α = 0.10.

```{r cars}
# Data: Mounting times (in minutes)
welles <- c(2, 4, 9, 3, 2)
atkins <- c(3, 7, 5, 8, 4, 3)

# Sample sizes
n_W <- length(welles)
n_A <- length(atkins)

# Sample means
mean_W <- mean(welles)
mean_A <- mean(atkins)

# Sample standard deviations
sd_W <- sd(welles)
sd_A <- sd(atkins)

# Store in a tibble
sample_stats <- tibble(
  `Method` = c("Welles", "Atkins"),
  `Sample Size (n)` = c(n_W, n_A),
  `Sample Mean` = c(mean_W, mean_A),
  `Sample Standard Deviation` = c(sd_W, sd_A)
)

sample_stats

```

Compute Pooled Variance

```{r}
# Compute pooled variance
sp2 <- ((n_W - 1) * sd_W^2 + (n_A - 1) * sd_A^2) / (n_W + n_A - 2)

# Store in a tibble
pooled_variance <- tibble(`Pooled Variance (s²_p)` = round(sp2, 3))

pooled_variance

```

Compute Test Statistic
```{r}
# Compute standard error
se <- sqrt(sp2 * (1/n_W + 1/n_A))

# Compute test statistic
t_test <- (mean_W - mean_A) / se

# Degrees of freedom
df <- n_W + n_A - 2

# Critical value for two-tailed t-test
t_critical <- qt(1 - 0.10 / 2, df)

# Store results in a tibble
test_statistic <- tibble(
  `Test Statistic (t)` = round(t_test, 3),
  `Degrees of Freedom` = df,
  `Critical Value (±t_critical)` = round(t_critical, 3)
)

test_statistic
```
Make the Decision
```{r}
# Decision rule
decision <- ifelse(abs(t_test) > t_critical, 
                   "Reject H₀: There is a significant difference in mounting times.", 
                   "Fail to Reject H₀: There is no significant difference in mounting times.")

# Store decision in a tibble
decision_table <- tibble(`Decision` = decision)

decision_table
```
Visualization: T-Test and Critical Regions
```{r}
# Create data for t-distribution
x_vals <- seq(-4, 4, length.out = 500)
y_vals <- dt(x_vals, df)

# Plot
ggplot(data.frame(x = x_vals, y = y_vals), aes(x = x, y = y)) +
  geom_line(color = "blue", size = 1.5) +  # T-distribution curve
  
  # Critical Regions (Shade Red)
  geom_area(data = subset(data.frame(x = x_vals, y = y_vals), x <= -t_critical),
            aes(y = y), fill = "darkred", alpha = 0.6) +
  geom_area(data = subset(data.frame(x = x_vals, y = y_vals), x >= t_critical),
            aes(y = y), fill = "darkred", alpha = 0.6) +
  
  # Test Statistic Line (Dashed Black Line)
  geom_vline(xintercept = t_test, color = "black", linetype = "dashed", size = 1.5) +  
  
  # Critical Value Lines (Dotted Red Lines)
  geom_vline(xintercept = -t_critical, color = "red", linetype = "dotted", size = 1.5) +  
  geom_vline(xintercept = t_critical, color = "red", linetype = "dotted", size = 1.5) +  
  
  # Labels for Test Statistic
  annotate("text", x = t_test, y = 0.05, label = paste0("Test Statistic = ", round(t_test, 2)),
           hjust = -0.1, size = 3, color = "black") +
  
  # Labels for Critical Values
  annotate("text", x = -t_critical, y = 0.05, label = paste0("-T Critical = ", round(-t_critical, 2)),
           hjust = -0.1, size = 2, color = "red") +
  annotate("text", x = t_critical, y = 0.05, label = paste0("+T Critical = ", round(t_critical, 2)),
           hjust = -0.1, size = 2, color = "red") +
  
  labs(
    title = "Two-Sample T-Test: Critical Region and Test Statistic",
    x = "T-Score",
    y = "Density"
  ) +
  theme_minimal(base_size = 15)

```

