---
title: "net migration"
author: "Kyaw Min Khaing"
date: "2025-02-07"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Create the initial tibble

```{r}
migration_data <- tibble(
  Region = c("Mandalay", "Bago", "Ayeyarwaddy", "Tanitharyi", "Myanmar"),
  Pop_2004 = c(7571, 5514, 7455, 1525, 54299),
  Pop_2008 = c(8216, 5879, 7858, 1665, 58377)
)
print(migration_data)
```

# Calculate the net migration rate

# Net Migration Rate Formula

The net migration rate $m$ for a region is calculated using the National Growth Rate method as follows:

$$
m = \left[ \frac{(P_{\text{t+n}} - P_t)}{P_t} - \frac{(P_{\text{T+n}} - P_T)}{P_T} \right] \times k
$$

Where: - $P_{\text{ret}}$ is the population at the end of the period. - $P_f$ is the population at the beginning of the period. - $P_i$ is the initial population of the region. - $P_T$ is the total population of the country. - $k$ is a constant (usually 100 for percentage).

## Mandalay

$$
m = \left[ \frac{(8216 - 7571)}{7571} - \frac{(58377 - 54299)}{54299} \right] \times 100 = 1.01\%
$$

## Bago

$$
m = \left[ \frac{(5879 - 5514)}{5514} - \frac{(58377 - 54299)}{54299} \right] \times 100 = -0.89\%
$$

## Ayeyarwaddy

$$
m = \left[ \frac{(7858 - 7455)}{7455} - \frac{(58377 - 54299)}{54299} \right] \times 100 = -2.1\%
$$

## Tanitharyi

$$
m = \left[ \frac{(1665 - 1525)}{1525} - \frac{(58377 - 54299)}{54299} \right] \times 100 = 1.68\%
$$

# Interpretation

According to the National Growth Rate method, during 2004 to 2008, there were net immigration in Mandalay Region and Tanitharyi Region. There were net out-migration in Ayeyarwaddy Region and Bago Region.

The occurrence of net in-migration is due to the better employment opportunities, better economic possibilities, facilities for education, housing, medical treatment, and higher standard of living in those areas.

```{r}
national_growth <- migration_data %>%
  filter(Region == "Myanmar") %>%
  mutate(
    Growth_Rate_National = (Pop_2008 - Pop_2004) / Pop_2004
  ) %>%
  pull(Growth_Rate_National) # နိုင်ငံတော် Growth Rate ကို သီးသန့်ထုတ်
print(national_growth)
```

```         
 3. Calculate Regional Migration Rate
```

```{r}
result <- migration_data %>%
  filter(Region != "Myanmar") %>% # နိုင်ငံတော်ဒေတာဖယ်ရန်
  mutate(
    Growth_Rate_Regional = (Pop_2008 - Pop_2004) / Pop_2004,
    Net_Migration_Rate = (Growth_Rate_Regional - national_growth) * 100
  )
print(result)
```

```         
# 4. Show Results
```

```{r}
result %>%
  select(Region, Net_Migration_Rate) %>%
  knitr::kable(digits = 2, caption = "Net Migration Rate (2004-2008)")

```

# Visualization using ggplot2

```{r}
ggplot(result, aes(x = reorder(Region, -Net_Migration_Rate), y = Net_Migration_Rate, fill = Region)) +
  geom_col() +
  geom_text(aes(label = sprintf("%.2f%%", Net_Migration_Rate)), vjust = -0.5) +
  labs(
    title = "Net Migration Rate by Region (2004-2008)",
    x = "Region",
    y = "Net Migration Rate (%)",
    caption = "Positive: In-migration | Negative: Out-migration"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
