---
title: "MAS122CDR and ASMR Calculation"
author: "Kyaw Min Khaing"
date: "2025-02-04"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Crude Death Rate (CDR) & Age-Specific Mortality Rate (ASMR)
1. Data Preparation

```{r cars}
# Create a tibble table with given data
death_data <- tibble(
  Age_Group = c("<5", "5-24", "25-44", "45-64", "65+"),
  Population = c(10900, 33840, 18840, 9840, 2760),
  Deaths = c(180, 75, 87, 158, 180)
)
print(death_data)
```
total_population
## Including Plots
```{r}
# Compute Total Population & Total Deaths
total_population <- sum(death_data$Population)
print(total_population)

```
Total_Deaths

```{r pressure, echo=FALSE}
total_deaths <- sum(death_data$Deaths)
print(total_deaths)

```
## **Age-Specific Mortality Rate (ASMR)**

The formula for the **Age-Specific Mortality Rate (ASMR)** is:

\[
M_x^n = ASMR = \frac{D_x^n}{P_x^n} \times 1000
\]

where:
- \( M_x^n \) = Mortality rate for age group \( x \)
- \( D_x^n \) = Number of deaths in age group \( x \)
- \( P_x^n \) = Population of age group \( x \)

For example, for the **age group 65+**:

\[
ASMR = \frac{180}{2760} \times 1000 = 65.22
\]
```{r}
# Compute ASMR for each age group
death_data <- death_data %>%
  mutate(ASMR = (Deaths / Population) * 1000)
print(death_data)
```
## **Crude Death Rate (CDR)**

The formula for the **Crude Death Rate (CDR)** is given by:

$$
CDR = \frac{D}{P} \times 1000
$$

where:
- \( D \) = Total number of deaths
- \( P \) = Total population

\[
CDR = \frac{680}{76180} \times 1000 = 8.93 \approx 9
\]
```{r}
# Compute CDR
CDR <- (total_deaths / total_population) * 1000
print(CDR)
```
Display the Results
```{r}
# Show Data Table with ASMR
death_data %>%
  mutate(ASMR = round(ASMR, 4)) %>%
  knitr::kable(caption = "Age-Specific Mortality Rate (ASMR) Table")
print(death_data)
```

```{r}
# Display CDR
cat("### **Crude Death Rate (CDR)**\n")
cat("CDR =", round(CDR, 0), "deaths per 1000 population\n")
```
Visualization of Age-Specific Mortality Rate (ASMR)
```{r}
ggplot(death_data, aes(x = Age_Group, y = ASMR, fill = Age_Group)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Age-Specific Mortality Rate (ASMR)",
    x = "Age Group",
    y = "ASMR (Deaths per 1000 Population)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(ASMR, 2)), vjust = -0.5)
```
Line Chart for ASMR Trends
```{r}
ggplot(death_data, aes(x = Age_Group, y = ASMR, group = 1)) +
  geom_line(color = "blue", size = 1) +
  geom_point(size = 2, color = "red") +
  labs(
    title = "ASMR Trends by Age Group",
    x = "Age Group",
    y = "ASMR (Deaths per 1000 Population)"
  ) +
  theme_minimal() +
  geom_text(aes(label = round(ASMR, 2)), vjust = -1, color = "black")
```
Visualization of Crude Death Rate (CDR)
```{r}
cdr_data <- tibble(
  Category = "Crude Death Rate",
  CDR = CDR
)

ggplot(cdr_data, aes(x = Category, y = CDR, fill = Category)) +
  geom_bar(stat = "identity", width = 0.4) +
  labs(
    title = "Crude Death Rate (CDR)",
    x = "Category",
    y = "CDR (Deaths per 1000 Population)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_text(aes(label = round(CDR, 2)), vjust = -0.5)
```

