---
title: "MAS112Incidence Rate of Pneumonia in Urban Area"
author: "Kyaw Min Khaing"
date: "2025-02-05"
output: word_document
---

## **Introduction**

This report calculates the incidence rate of pneumonia over a 10-year period among a population of **118,540** people in an urban area.

### **Formula for Incidence Rate**

$$
Incidence\ Rate = \left( \frac{\text{New Cases}(t)}{\text{Population}(t)} \right) \times 10,000
$$

```{r setup, include=FALSE}
library(tidyverse)
```

## \# Given Data

```{r cars}
pneumonia_data <- tibble(
  Year = 2000:2009,  # 10-year period
  Total_Population = rep(118540, 10),
  New_Cases = rep(322/10, 10),  # Average cases per year
  Incidence_Rate = (322/10) / 118540 * 10000  # Incidence Rate Calculation
)

# Display the tibble table
pneumonia_data
```

## Visualization of Pneumonia Incidence Rate

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ggplot(pneumonia_data, aes(x = Year, y = New_Cases)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = round(New_Cases, 1)), vjust = -0.5, size = 5) +
  labs(title = "Incidence of Pneumonia Cases Over 10 Years",
       x = "Year",
       y = "New Pneumonia Cases per Year") +
  theme_minimal()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
