---
title: "Kruskal Wallis Test"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2025-01-27"
---
date: "2025-01-27"
Presented by Group 1 students 
1. Ma Cho Cho Yee (MAS 1 - 8)
2. Mg Kaung Myat Thu (MAS 1 - 13)
3. Mg Kyaw Min Khaing (MAS 1 - 19)
4. Mg Kyaw Swar Tun (MAS 1 - 20)
5. Mg Nay Aung (MAS 1 - 30)
6. Mg Nay Myo Aung (MAS 1 - 31)
7. Mg Soe Eain Htet (MAS 1 - 34)
8. Mg Thet Paing Htet (MAS 1 - 39)
9. Mg Wai Linn Htet Thwin (MAS 1 - 43)
10 Mg Yan Naing Soe (MAS 1 - 51)
---

Kruskal-Wallis H-test statistic using the formula: 
$$
H = \frac{12}{N(N+1)} \sum \frac{R_i^2}{n_i} - 3(N+1)$$

Where: - $N$ is the total number of observations. - $R_i$ is the rank sum for each group. - $n_i$ is the number of observations in each group.

Assumption-

1.  non-parametric statistical test
2.  three or more independent groups
3.  if there are statistically significant differences between them ,it is an extension of the Mann_Whitney U test which is used for comparing two groups.

In this 1974 Motor Trend US magazine , is there a significant difference between Cylinder Count (cyl) and Horsepower (hp)? Assess the normality assumption for the data? Use the Kruskal-Wallis Test to evaluate whether there are statistically significant differences in Horsepower across the groups (4-cylinder, 6-cylinder, 8-cylinder). Present the test results using a boxplot and appropriately annotate the p-value in your visualization.

## State the Null and Research Hypotheses

### Alternative Hypothesis (H1):is HO: θL = θM = θH (All groups have the same median).

### The research hypothesis is HA: There is a tendency for self-conidence to rank systematically higher or lower for at least one level of social interaction when compared with the other levels. (at least one group differs in its median).

### Set the Level of Risk (or the Level of Signiicance) Associated with the Null Hypothesis α = 0.05, 95% chance that any observed statistical difference will be real and not due to chance

### Load necessary packages

```{r}
library(pacman)
p_load(rmarkdown,officer,knitr,kableExtra,latex2exp,xtable,flextable,tidyverse, rstatix, ggsignif)
```

### Convert knit format

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,    
  warning = FALSE, 
  message = FALSE, 
  fig.width = 6,   
  fig.height = 4   
)
```

### check my data type

```{r}
mtcars <- as.tibble(mtcars)
glimpse(mtcars)
```

### show my data

```{r}
mtcars <- as.tibble(mtcars)
kable(mtcars, caption = "mtcars Dataset")
```

### Convert to long format(tidy) for analysis and data manipulation

```{r}
data_long <- mtcars  |> 
  select(hp, cyl)  |> 
  mutate(cyl = as.factor(cyl))  |> 
  pivot_longer(cols = hp, names_to = "Metric", values_to = "Value") 
kable(data_long, caption = "data cleaning and wrangling")
```

### calculate rank and handling Ties for the Test Statistic doing Ties Correction (C_H)

```{r}
ranked_data <- data_long |> 
  mutate(Rank = rank(Value), 
         Tie_Case = duplicated(Value) | duplicated(Value, fromLast = TRUE)) |> 
  arrange(Rank)  # Rank အလိုက် စဉ်သည်

kable(ranked_data, caption = "sorting,Ranking and Tie")

```

### Total Rank Variability for hp

```{r}
rank_sums <- ranked_data |> 
  group_by(cyl) |>   # Ensure cyl exists
  summarise(
    n = n(),
    Rank_Sum = sum(Rank)
  )

# Print rank sums
kable(rank_sums, caption = "total Rank")

```

### total number of observation

```{r}
N <- nrow(ranked_data)

cat("Total number of observation (N):", N, "\n")
```

### number of category

```{r}
k <- n_distinct(ranked_data$cyl)
cat(" Number of category (N):", k, "\n")
```

## Compute the Kruskal-Wallis H-test statistic using the formula:

$$
H = \frac{12}{N(N+1)} \sum \frac{R_i^2}{n_i} - 3(N+1)$$

Where: - $N$ is the total number of observations. - $R_i$ is the rank sum for each group. - $n_i$ is the number of observations in each group.

```{r}
df <- k - 1

H <- 12 / (N * (N + 1)) * sum((rank_sums$Rank_Sum^2) / rank_sums$n) - 3 * (N + 1)
p_value <- pchisq(H, df, lower.tail = FALSE)

cat("Kruskal-Wallis H statistic:", round(H, 4), "\n")
cat("Degrees of Freedom:", df, "\n")
cat("p-value:", round(p_value, 6), "\n")
if (p_value < 0.05) {
  cat("The result is significant (p < 0.05). At least one group differs.\n")
} else {
  cat("The result is not significant (p >= 0.05). Groups are similar.\n")
}
```

## Compute ties table based on Tie_Case

```{r}
ties_table <- ranked_data |> 
  filter(Tie_Case) |> 
  count(Rank)

# Print the updated ties table
kable(ties_table, caption = "Number of Tie")

```

# Compute C_H correction factor

## To Compute the Test Statistic doing Ties Correction (C_H)

The ties correction is applied to adjust for tied ranks. The formula for the ties correction $C_H$ is:

$$
C_H = 1 - \frac{\sum{(T^3 - Ti)}}{N^3 - N}
$$

Where: - $T$ is the number of tied ranks in each group of ties. - $N$ is the total number of observations.

```{r}
# Compute C_H correction factor
C_H <- 1 - sum((ties_table$n^3 - ties_table$n) / ((nrow(ranked_data)^3) - nrow(ranked_data)))

# Print C_H
cat("Ties Correction Factor (C_H):", round(C_H, 4), "\n")
```

Calculate H statistic

```{r}
H_corrected <- H / C_H 
cat("Corrected H Statistic:", round(H_corrected, 4), "\n")
```

# **Pairwise Wilcoxon Rank Sum Test with Bonferroni Correction**

```{r}
pairwise_results <- ranked_data |> 
  pairwise_wilcox_test(Value ~ cyl, p.adjust.method = "bonferroni") |> 
  mutate(p = round(p, 6), p.adj = round(p.adj, 3))

kable(pairwise_results, caption = "pairwise results")

```

# Visualizing pairwise comparisons with exact p-values with heatmap

```{r}
filtered_results <- pairwise_results |> 
  filter((group1 == "4" & group2 == "6") | 
         (group1 == "4" & group2 == "8") | 
         (group1 == "6" & group2 == "8"))
ggplot(filtered_results, aes(x = group1, y = group2, fill = p.adj)) +
  geom_tile(color = "white") +  # Heatmap style
  geom_text(aes(label = round(p.adj, 3)), color = "black", size = 5) +  # Adjusted p-values (3 digits)
  scale_fill_gradient(low = "blue", high = "red") +  # Color scale for p-values
  theme_minimal() +
  labs(title = "Pairwise Comparisons (Adjusted p-values)", 
       x = "Group 1 (cyl)", 
       y = "Group 2 (cyl)", 
       fill = "Adjusted p-value") +
  theme_minimal() 
```

visturalization boxplot

```{r}
rank_plot <- ggplot(data_long , aes(x = cyl, y = Value, fill = cyl)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8, alpha = 0.7) +  # Transparent boxplot
  geom_jitter(aes(color = cyl), width = 0.2, size = 2, alpha = 0.7) +  # Jitter points
  labs(
    title = "Rank Distribution of HP by Cylinders",
    x = "Number of Cylinders",
    y = "Value of hp"
  ) +
  theme_minimal() +
  geom_signif(
    comparisons = list(
      c("4", "6"),
      c("4", "8"),
      c("6", "8")
    ),
    map_signif_level = TRUE,
    test = "wilcox.test",
    step_increase = 0.1
  ) +
  scale_color_manual(values = c("4" = "blue", "6" = "green", "8" = "purple")) +  # Custom colors
  theme(legend.position = "none")  # Hide legend

print(rank_plot)
```

# Perform Kruskal-Wallis Test

```{r}
rstatix_in_kruskal_result <- data_long |> 
  kruskal_test(Value ~ cyl)
kable(rstatix_in_kruskal_result, caption = "Rstatix in kruskal result")
```

# Perform pairwise Wilcoxon comparisons with Bonferroni correction

```{r}
rstatix_in_pairwise_results <- data_long |> 
  pairwise_wilcox_test(Value ~ cyl, p.adjust.method = "bonferroni") |> 
  mutate(p = round(p, 6), p.adj = round(p.adj, 3))
kable(rstatix_in_pairwise_results, caption = "rstatix_in_pairwise_results")
```
