---
title: "Sample Point-Biserial Correlation (Small Data Samples)"
author: "Kyaw Min Khaing"
date: "2025-01-29"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Introduction

This analysis applies the **Point-Biserial Correlation** to examine the relationship between **Gender (binary variable: Male/Female)** and **Posttest Scores** for a small dataset.

## Step 1: State the Hypotheses

-   **Null Hypothesis (H₀):** There is no significant correlation between gender and posttest scores.
-   **Alternative Hypothesis (H₁):** There is a significant correlation between gender and posttest scores.

## Step 2: Set the Level of Significance

We set **α = 0.05** as the significance level.

## Step 3: Load and Prepare Data

```{r cars}
data <- tibble(
  Participant = 1:17,
  Gender = c(rep("M", 8), rep("F", 9)),
  Posttest_Score = c(7, 19, 8, 10, 7, 15, 6, 13, 14, 11, 18, 23, 17, 20, 14, 24, 22)
  ) %>% mutate(Gender_Binary = ifelse(Gender == "M", 0, 1)) 

print(data)
```

```{r}
n <- nrow(data)
```

## Step 4: Compute Point-Biserial Correlation

The formula for Point-Biserial Correlation is:

$$
 r_{pb} = \frac{M_1 - M_0}{s} \times \sqrt{p_0 \times p_1}
$$

where: - $M_1$ is the mean of the continuous variable for one group (e.g., females).
- $M_0$ is the mean of the continuous variable for the other group (e.g., males).
- $s$ is the standard deviation of the continuous variable.
- $p_1$ is the proportion of one group.
- $p_0$ is the proportion of the other group.

#### The mean females’ posttest score was

```{r}
M_1 <- mean(data %>% filter(Gender_Binary == 1) %>% pull(Posttest_Score))
glimpse(M_1)
```

#### The mean males’ posttest score was

```{r}
M_0 <- mean(data %>% filter(Gender_Binary == 0) %>% pull(Posttest_Score))
glimpse(M_0 )
```

#### standard deviation for the interval data

```{r}
s <- sd(data$Posttest_Score)
glimpse(s)
```

number of males and females

```{r}
n <- nrow(data)
glimpse(n )
```

number of females

```{r}
n_1 <- sum(data$Gender_Binary == 1)
glimpse(n_1)
```

#### Proportion of females

```{r}
p_1 <- n_1 / n
glimpse(p_1)
```

number of males

```{r}
n_0 <- sum(data$Gender_Binary == 0)
glimpse(n_0)
```

#### Proportion of males

```{r}
p_0 <- n_0 / n
glimpse(p_0)
```

## The formula for Point-Biserial Correlation is:

$$
 r_{pb} = \frac{M_1 - M_0}{s} \times \sqrt{p_0 \times p_1}
$$

where: - $M_1$ is the mean of the continuous variable for one group (e.g., females).
- $M_0$ is the mean of the continuous variable for the other group (e.g., males).
- $s$ is the standard deviation of the continuous variable.
- $p_1$ is the proportion of one group.
- $p_0$ is the proportion of the other group.

#### the point-biserial correlation coeficient using the values computed 
 
```{r}
r_pb <- (abs((M_0 - M_1) / s) * sqrt(p_0 * p_1)) 
glimpse(r_pb)
```
```{r}
cor_test_result <- data %>%
  cor_test(Gender_Binary, Posttest_Score, method = "pearson")
# Result ကို print လုပ်ပါ
print(cor_test_result)
```


## Step 5: Determine the Critical Value

For **n = 17** and **α = 0.05 (two-tailed test)**, the critical value from the **t-distribution table** is **±0.482**.

## Step 6: Compare the Obtained Value with the Critical Value

```{r}
if (abs(r_pb) >= 0.482) {
  cat("The correlation is statistically significant. We reject H₀ and conclude there is a significant correlation.\n")
} else {
  cat("The correlation is not statistically significant. We fail to reject H₀.\n")
}
```

## Step 7: Interpret the Results

If **\|r_pb\| ≥ 0.482**, the result is significant.

## Step 8: Reporting the Results

> *A Point-Biserial correlation was conducted to determine the relationship between gender and posttest scores. The analysis yielded r_pb = `r round(r_pb, 4)`. Since the obtained r_pb is `r ifelse(abs(r_pb) >= 0.482, "greater", "less")` than the critical value of 0.482, we `r ifelse(abs(r_pb) >= 0.482, "reject", "fail to reject")` the null hypothesis. This suggests that there `r ifelse(abs(r_pb) >= 0.482, "is", "is no")` significant correlation between the two variables.*

## Step 9: Data Visualization

To better understand the distribution of posttest scores by gender, we visualize it using `ggplot2`.

```{r}


ggplot(data, aes(x = Gender_Binary, y = Posttest_Score)) +
  geom_point(size = 4, color = "blue") +  # Data points များကို ပြသပါ
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Trend line ထည့်ပါ
  labs(
    title = "Scatterplot of Gender (Binary) vs Posttest Score",
    x = "Gender (0 = Male, 1 = Female)",
    y = "Posttest Score"
  ) +
  theme_minimal() 
```
```{r}
data_long <- data %>%
  pivot_longer(cols = Posttest_Score, names_to = "Variable", values_to = "Score")
ggplot(data_long, aes(x = Gender, y = Score, fill = Gender)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +
  labs(
    title = "Distribution of Posttest Scores by Gender",
    x = "Gender",
    y = "Posttest Score"
  ) +
  theme_minimal()
```

