---
title: "Sample Runs Test (Large Data Samples)"
author: "Kyaw Min Khaing"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Sample Runs Test (Large Data Samples)

### 1. Hypothesis Statements
- Null Hypothesis \(H_0\): The sequence is random.
- Alternative Hypothesis \(H_A\): The sequence is not random.

### 2. Define Data

```{r}
# Given data
# Given data
n1 <- 23  # Number of occurrences of category 1 (e.g., Female)
n2 <- 14  # Number of occurrences of category 2 (e.g., Male)
R <- 17   # Number of runs observed in the sequence

# Total sample size
n <- n1 + n2

# Create a tibble table for input values
input_data <- tibble(
  "Parameter" = c("Number of Female (n1)", "Number of male (n2)", 
                  "Total Sample Size (n)", "Observed Runs (R)"),
  "Value" = c(n1, n2, n, R)
)

# Display input table
kable(input_data, caption = "number of Female and male")
```

### 3. Compute Number of Runs \( R \)
$$
\bar{X}_R = \frac{2n_1n_2}{n_1 + n_2} + 1
$$

```{r}
# Compute mean of runs
X_R <- (2 * n1 * n2) / (n1 + n2) + 1
X_R
```

### 4. Compute Mean and Standard Deviation

$$
s_R = \sqrt{\frac{2n_1n_2(2n_1n_2 - n_1 - n_2)}{(n_1 + n_2)^2 (n_1 + n_2 - 1)}}
$$


```{r}
# Compute standard deviation of runs
s_R <- sqrt((2 * n1 * n2 * (2 * n1 * n2 - n1 - n2)) / ((n1 + n2)^2 * (n1 + n2 - 1)))

print(s_R)
```

### 5. Compute z-score
$$
Z = \frac{R+ h- \bar{x}_R}{s_R}
$$


```{r}
h <- 0.5 
Z <- (R + h - X_R) / s_R
print(Z)
```

### 6. Decision-Making

```{r}
alpha <- 0.05  # Significance level
critical_value <- qnorm(1 - alpha / 2)  # Two-tailed test

# Decision based on Z-score
decision <- ifelse(abs(Z) > critical_value, 
                   "Reject H0: The sequence is not random.", 
                   "Fail to reject H0: The sequence appears random.")

# Create a tibble table for decision making
decision_table <- tibble(
  "Parameter" = c("Significance Level (α)", "Critical Value (±Z_α/2)", "Decision"),
  "Value" = c(alpha, round(critical_value, 3), decision)
)

# Display decision table
kable(decision_table, caption = "decision")
```

### 7. Visualization

```{r}
ggplot(data.frame(x = c(-4, 4)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), color = "blue") +
  geom_vline(xintercept = Z, color = "red", linetype = "dashed") +
  geom_vline(xintercept = critical_value, color = "black", linetype = "dotted") +
  geom_vline(xintercept = -critical_value, color = "black", linetype = "dotted") +
  labs(title = "Z-Score in Standard Normal Distribution",
       x = "Z-Score",
       y = "Density") +
  theme_minimal()
```

### 8. Interpretation
Based on the computed z-score and critical values, we conclude whether the sequence is random or not.
