---
title: "Sample_Point-Biserial_Correlation_(Large_Data _Samples)"
author: "Kyaw Min Khaing"
date: "2025-01-30"
output:
  word_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
library(tidyverse)
```

## R Markdown

Step 1: State the Hypotheses

Null Hypothesis (H₀): There is no significant correlation between gender and posttest scores.

Alternative Hypothesis (H₁): There is a significant correlation between gender and posttest scores.

Step 3: Load and Prepare Data

```{r cars}
data <- tibble(
  Participant = 1:26,
  Gender = c(rep("M", 14), rep("F", 12)),
  Posttest_Score = c(6, 15, 8, 10, 6, 12, 7, 13, 13, 10, 18, 23, 17, 20, 14, 26, 14, 11, 29, 20, 15, 18, 9, 14, 21, 22)
  ) %>% 
  mutate(Gender_Binary = ifelse(Gender == "M", 0, 1)) 

print(data)
```

## The formula for Point-Biserial Correlation is:

$$
 r_{pb} = \frac{M_1 - M_0}{s} \times \sqrt{p_0 \times p_1}
$$

where: - $M_1$ is the mean of the continuous variable for one group (e.g., females).
- $M_0$ is the mean of the continuous variable for the other group (e.g., males).
- $s$ is the standard deviation of the continuous variable.
- $p_1$ is the proportion of one group.
- $p_0$ is the proportion of the other group.
#### means for female

```{r pressure, echo=FALSE}
M_1 <- mean(data %>% filter(Gender_Binary == 1) %>% pull(Posttest_Score))
cat("means for female:", round(M_1, 4), "\n")
```

#### means for male

```{r}
M_0 <- mean(data %>% filter(Gender_Binary == 0) %>% pull(Posttest_Score))
cat("means for male:", round(M_0, 4), "\n")
```

#### Compute standard deviation

```{r}
s <- sd(data$Posttest_Score)
cat("standard deviation for the interval:", round(M_0, 4), "\n")
```

#### Total sample size

```{r}
n <- nrow(data)
cat("total number of people:", round(n, 4), "\n")
```

## proportions for female

```{r}
p_1 <- sum(data$Gender_Binary == 1) / n
cat("proportions for female", round(p_1 , 4), "\n")
```

## proportions for male

```{r}
p_0 <- sum(data$Gender_Binary == 0) / n
cat("proportions for male", round(p_0 , 4), "\n")
```

## Step 4: Compute Point-Biserial Correlation

```{r}
r_pb <- (abs((M_0- M_1) / s) * sqrt(p_0 * p_1))
cat("Point-Biserial Correlation (r_pb):", round(r_pb, 4), "\n")
```

\## Step 5: Determine the Critical Value

Using an approximation to the normal distribution:

$$
 z^* = r_{pb} \times \sqrt{n - 1}
$$

```{r}
Z_star<- (r_pb * sqrt(n - 1))
cat("critical value (Z_star):", round(Z_star, 4), "\n")
```

From Table B.1 in Appendix B, for a two-tailed test with $\alpha = 0.05$, we must not reject $H_0$ if:

$$
-1.96 \leq z^* \leq 1.96
$$ \## Step 6: Compare the Obtained Value with the Critical Value

```{r}
if ((Z_star) > 1.96) {
  conclusion <- "The correlation is statistically significant. We reject H₀ and conclude there is a significant correlation."
} else {
  conclusion <- "The correlation is not statistically significant. We fail to reject H₀."
}

print(conclusion)
```

## Step 7: Interpret the Results

Since $z^* = 2.055$ is greater than the critical value of 1.96, we reject the null hypothesis, suggesting that the correlation between gender and visual detail recognition is real.
This suggests a **moderate but weak correlation** between gender and visual detail recognition.

## Step 8: Reporting the Results

> *A Point-Biserial correlation was conducted to determine the relationship between gender and visual detail recognition. The analysis yielded* $r_{pb} = `r round(r_pb, 4)`$.
> The computed z-score was $z^* = `r round(z_star, 3)`$.
> Since $z^*$ is `r ifelse(abs(z_star) > 1.96, "greater", "less")` than the critical value of 1.96, we `r ifelse(abs(z_star) > 1.96, "reject", "fail to reject")` the null hypothesis.
> This suggests that there `r ifelse(abs(z_star) > 1.96, "is", "is no")` significant correlation between the two variables.

A study comparing male and female ability to recognize and remember visual details involved **14 males (nM = 14)** and **12 females (nF = 12)**.
The mean recognition score for **males was 12.71**, whereas the mean recognition score for **females was 17.75**.
The correlation coefficient $r_{pb}$ was **0.411**, and the results were statistically significant **(p \< 0.05)**, indicating a **moderate relationship between gender and visual detail recognition**.

## Step 9: Data Visualization

```{r}
data_long <- data %>%
  pivot_longer(cols = Posttest_Score, names_to = "Variable", values_to = "Score")

ggplot(data_long, aes(x = Gender, y = Score, fill = Gender)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7) +
  labs(
    title = "Distribution of Visual Detail Recognition Scores by Gender",
    x = "Gender",
    y = "Posttest_Score"
  ) +
  theme_minimal()
```
