---
title: "Sample_Biserial_Correlation_(Small_Data_Samples)"
author: "Kyaw Min Khaing"
date: "2025-02-02"
output: html_document
---

necessary package

```{r}
library(tidyverse,knitr)
```

preapare my data

```{r}
data <- tibble(
  Participant = 1:16,
  Exam_Performance = c(rep("F", 5), rep("P", 11)),  
  GPA = c(3.5, 3.4, 3.3, 3.2, 3.6, 
          4.0, 3.6, 4.0, 4.0, 3.8, 
          3.9, 3.9, 4.0, 3.8, 3.5, 3.6)
) %>% mutate(Exam_Binary = ifelse(Exam_Performance == "P", 1, 0)) 

kable(data, caption = "Exam performance")
```
## State the Null and Research Hypothesis 
HO: ρb = 0
HA: ρb ≠ 0

## Set the Level of Risk (or the Level of Signiicance) Associated with the Null Hypothesis

The level of risk, also called an alpha (α), is frequently set at 0.05. We will use α = 0.05 in our example. In other words, there is a 95% chance that any observed statistical difference will be real and not due to chance.

```{r}
alpha <- 0.05
```

## Compute the Test Statistic 


```{r cars}
# Compute means for each group
M_1 <- mean(data %>% filter(Exam_Binary == 1) %>% pull(GPA))
cat("means for pass:", round(M_1, 4), "\n")
```

```{r}
M_0 <- mean(data %>% filter(Exam_Binary == 0) %>% pull(GPA))
cat("means for fail:", round(M_0, 4), "\n")

```

```{r}
# Compute standard deviation
s <- sd(data$GPA)
cat("standard deviation for data:", round(s, 4), "\n")
```

```{r}
# Compute proportions
n_1 <- sum(data$Exam_Binary == 1)
cat("number for pass:", round(n_1, 4), "\n")
```


```{r}
n_0 <- sum(data$Exam_Binary == 0)
cat("number for fail:", round(n_0, 4), "\n")
```


```{r}
n <- n_1 + n_0
cat("number for pass and fail:", round(n, 4), "\n")
```


```{r}
p_1 <- n_1 / n
cat("proportion of pass:", round(p_1, 4), "\n")

```


```{r}
p_0 <- n_0 / n
cat("proportion of fail:", round(p_0, 4), "\n")
```
Now, determine the height of the unit normal curve ordinate, y, at the point dividing 
Pp and Pq. We could reference the table of values for the normal distribution, such 
as Table B.1 in Appendix B, to ind y. However, we will compute the value. Using 
Table B.1 also provides the z-score at the point dividing Pp and Pq, z = 0.49:

$$
y = \frac{1}{\sqrt{2\pi}} e^{-z^2/2}
$$

```{r}
# Compute r_b
z <- 0.49
y <- (1 / sqrt(2 * pi)) * exp(-0.5 * z^2)
print(y)
```


```{r}
r_b <- abs(((M_1 - M_0) / s) * (p_1 * p_0) / y)
print(r_b)
```
Determine the Value Needed for Rejection of the Null Hypothesis 
Using the Appropriate Table of Critical Values for the Particular Statistic Table B.8 in Appendix B lists critical values for the Pearson product-moment 
correlation coeficient. The table requires the degrees of freedom and df = n − 2. 
In this study, n = 16 and df = 16 − 2. Therefore, df = 14. Since we are conducting 
a two-tailed test and α = 0.05, the critical value is 0.497.
7.5.6.6 Compare the Obtained Value with the Critical Value The critical 
value for rejecting the null hypothesis is 0.497 and the obtained value is |rb| = 0.972. 
If the critical value is less than or equal to the obtained value, we must reject the 
null hypothesis. If instead, the critical value is greater than the obtained value, we 
must not reject the null hypothesis. Since the critical value is less than the absolute 
value of the obtained value, we reject the null hypothesis.
7.5.6.7 Interpret the Results We rejected the null hypothesis, suggesting that 
there is a signiicant and very strong correlation between student GPA and comprehensive exam performance.
