---
title: "Sample Fisher Exact Test"
author: "Kyaw Min Khaing"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Step 1: State the Null and Research Hypotheses
Null Hypothesis: Gender and Attitude are independent (no association).
Alternative Hypothesis: Gender and Attitude are not independent (there is an association).
Step 2: Set the Level of Significance
The significance level (α) is set at 0.05 (5%).

```{r cars}
# Level of significance
alpha <- 0.05
```

## preparation

```{r pressure, echo=FALSE}
# Create dataset using tibble
data <- tibble(
  Participant = 1:11,
  Gender = c("Male", "Male", "Male", "Male", "Male", 
             "Female", "Female", "Female", "Female", "Female", "Female"),
  Score = c(+30, +14, -21, +22, +9, -22, -13, -20, -7, +19, -31),
  Attitude = c("+", "+", "-", "+", "+", "-", "-", "-", "-", "+", "-")
)
kable(data, caption = "number of Female and male")

```

Construct the Contingency Table
We summarize the data to create a 2x2 contingency table.
```{r}
# Create a contingency table
contingency_table <- data %>%
  group_by(Gender, Attitude) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  pivot_wider(names_from = Attitude, values_from = Frequency, values_fill = 0)

# Rename columns for clarity
contingency_table <- contingency_table %>%
  rename(Positive = `+`, Negative = `-`)

# Display contingency table
contingency_table
```
 Step 5: Compute the Fisher Exact Test Statistic
The Fisher Exact Test is based on combinations and the hypergeometric probability formula:
```{r}
# Extract values from the contingency table
a <- contingency_table$Positive[1]  # Male Positive
b <- contingency_table$Negative[1]  # Male Negative
c <- contingency_table$Positive[2]  # Female Positive
d <- contingency_table$Negative[2]  # Female Negative

# Compute total sample size
n <- a + b + c + d

# Compute Fisher's exact probability using factorials
factorial_fn <- function(x) {
  if (x == 0) return(1) else return(prod(1:x))
}

# Compute probability
p_value <- (factorial_fn(a + c) / (factorial_fn(a) * factorial_fn(c))) *
           (factorial_fn(b + d) / (factorial_fn(b) * factorial_fn(d))) /
           (factorial_fn(n) / (factorial_fn(a + b) * factorial_fn(c + d)))

# Display computed p-value
glimpse(p_value)
```
Determine the Critical Value
Since Fisher's Exact Test is non-parametric, we compare the computed p-value with the significance level α = 0.05.
```{r}
# Compare p-value with alpha
if (p_value < alpha) {
  result <- "Reject the null hypothesis (significant association between Gender and Attitude)."
} else {
  result <- "Fail to reject the null hypothesis (no significant association between Gender and Attitude)."
}

# Display re
result
```
Interpretation and Conclusion
```{r}
cat("The computed p-value is:", p_value, "\n")
cat("Significance level (α):", alpha, "\n")
cat("Conclusion:", result, "\n")
```

 