---
title: "Sample_χ 2Goodness-of-Fit Test (Category Frequencies  Not Equal)"
author: "Kyaw Min Khaing"
date: "2025-02-02"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Step 1: Set the Level of Significance
The level of significance (α) is typically set at 0.05 (5%). This represents a 5% risk of rejecting the null hypothesis when it is actually true.

```{r}
alpha <- 0.05
```

Define the Observed and Expected Frequencies
```{r}
# Observed frequencies for the programs
data <- tibble(
  Program = c("Program 1", "Program 2", "Program 3"),
  Observed = c(110, 55, 85),
  Expected_Proportion = c(0.32, 0.22, 0.45)  # Hypothesized proportions
)
# Compute total observations
total_observed <- sum(data$Observed)
# Compute expected frequencies
data <- data %>%
  mutate(Expected = Expected_Proportion * total_observed)
# Display dataset
print(data)
```

Compute the Chi-Square Test Statistic
```{r}
# Compute chi-square test statistic
chi_square_statistic <- data %>%
  mutate(Chi_Square_Term = (Observed - Expected)^2 / Expected) %>%
  summarise(Chi_Square = sum(Chi_Square_Term)) %>%
  pull(Chi_Square)

# Display the computed chi-square statistic
glimpse(chi_square_statistic)
```
Step 4: Determine the Degrees of Freedom

```{r}
# Compute degrees of freedom
df <- nrow(data) - 1

# Display degrees of freedom
df

```
Find the Critical Value
```{r}
# Compute the critical value
critical_value <- qchisq(1 - alpha, df)

# Display the critical value
critical_value
```
Compare Test Statistic to Critical Value
```{r}
# Decision based on comparison
if (chi_square_statistic > critical_value) {
  result <- "Reject the null hypothesis"
} else {
  result <- "Fail to reject the null hypothesis"
}

# Display decision
result

```
Conclusion
```{r}
# Display conclusion
cat("The computed chi-square statistic is:", chi_square_statistic, "\n")
cat("The critical value is:", critical_value, "\n")
cat("Conclusion:", result, "\n")

```

