---
title: "Sample Runs Test for Small Data Samples"
author: "Kyaw Min Khaing"
date: "2025-02-03"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
```

Step 1: State the Null and Research Hypotheses
Null Hypothesis : The sequence of categories (M/F) is random.
Alternative Hypothesis : The sequence of categories (M/F) is not random (suggesting clustering or patterns).
Step 2: Set the Level of Significance
The significance level (α) is set at 0.05 (5%).

```{r cars}
# Level of significance
alpha <- 0.05
```

Step 3: Create the Dataset
We define the sequence of Male (M) and Female (F) groupings and represent them as a vector.

```{r pressure, echo=FALSE}
# Define sequence dataset
data <- tibble(
  Position = 1:13,
  Category = c("M", "FF", "M", "F", "M", "F", "M", "FF", "M", "FFF", "MM", "F", "MMM")
)

# Display dataset
data

```

Step 4: Count the Total Runs in the Sequence
A run is defined as a sequence of the same category (M or F) occurring consecutively.
```{r}
# Convert sequence into binary format for easy computation
binary_seq <- data$Category %>% str_extract("^M|^F")

# Count number of runs
num_runs <- sum(binary_seq[-1] != binary_seq[-length(binary_seq)]) + 1

# Display number of runs
num_runs

```
Compute the Expected Number of Runs and Standard Deviation

The expected number of runs (\(E(R)\)) and the standard deviation (\(\sigma_R\)) are given by:

\[
E(R) = \frac{2n_1 n_2}{n} + 1
\]

\[
\sigma_R = \sqrt{\frac{2n_1 n_2(2n_1 n_2 - n)}{n^2(n-1)}}
\]

where:
- \( n_1 \) = number of **M** (Male occurrences)
- \( n_2 \) = number of **F** (Female occurrences)
- \( n = n_1 + n_2 \)
```{r}
# Count Males (M) and Females (F)
n1 <- sum(str_detect(data$Category, "M"))
n2 <- sum(str_detect(data$Category, "F"))
n <- n1 + n2

# Compute expected runs and standard deviation
E_R <- (2 * n1 * n2) / n + 1
sigma_R <- sqrt((2 * n1 * n2 * (2 * n1 * n2 - n)) / (n^2 * (n - 1)))

# Display expected runs and standard deviation
E_R
sigma_R


```
Step 6: Compute the Z-score
```{r}
# Compute Z-score
Z_score <- (num_runs - E_R) / sigma_R

# Display Z-score
Z_score
```
Step 7: Determine the Critical Value
Using a standard normal distribution table, the critical value for α = 0.05 (two-tailed test)

```{r}
# Compute critical value
critical_value <- qnorm(1 - alpha/2)
# Display critical value
critical_value
```
Compare the Obtained Value with the Critical Value
If |Z-score| > critical value, we reject the null hypothesis.


```{r}
# Decision based on comparison
if (abs(Z_score) > critical_value) {
  result <- "Reject the null hypothesis (the sequence is not random)."
} else {
  result <- "Fail to reject the null hypothesis (the sequence is random)."
}

# Display decision
result

```
 Conclusion
```{r}
cat("Number of Runs (R):", num_runs, "\n")
cat("Expected Runs (E(R)):", E_R, "\n")
cat("Standard Deviation (σR):", sigma_R, "\n")
cat("Computed Z-score:", Z_score, "\n")
cat("Critical Value (two-tailed):", critical_value, "\n")
cat("Conclusion:", result, "\n")
```
 
